Formalizing Stack Safety as a Security Property

### 2023-02-03 submission5/core.tex
Halfway through figures

\apt{Add brief case for having formal guarantees.}
Of these mechanisms, many are fundamentally ill-suited for offering formal guarantees,
as they aim to impede attackers but cannot provide universal protection. Shadow stacks,
for instance, aim to ``restrict the flexibility available in creating gadget chains''
\cite{Shanbhogue+19}, not to categorically
rule out attacks. Other mechanisms such as Softbound\cite{NagarakatteZMZ09} and Code-pointer
integrity\cite{Kuznetsov+14} are stronger, but most do not give formal guarantees.
Indeed, the secure calling conventions of Skorstengaard et al. \cite{SkorstengaardSTKJFP} and
Georges et al. \cite{Georges22:TempsDesCerises} is so far the sole line of work
to offer stack safety formalizations.

@@

\section{Key Ideas}
\label{sec:ideas}

\apt{Not a good place for this.}
The paper is structured as follows. In \cref{sec:example}, we will walk through a function
call in a simple example machine and discuss how each of our properties applies to it,
informally. In the process we discuss some motivation for the properties, from a security
perspective. We consider a simple attacker, and discuss the harm that it can do to
its caller. This perspective is essential to understanding how lazy enforcement can
be safe: even if an attacker can overwrite some memory, if it cannot accomplish
subvert the observable behavior of the caller, the system is safe.

In \cref{sec:machine} we formalize the machine model, 
its {\em security semantics}, and the stack safety properties built on these. 
\Cref{sec:testing} describes the testing framework, and
\cref{sec:relwork,sec:future} discuss related and future work.
The remainder of this section introduces ideas that will be important throughout.

\paragraph*{Building a Property}

@@

\newcommand{\textbox}[1][]{\tikz \filldraw[fill=lightgray] (0,0) rectangle (30px,30px) node[pos=.5]{#1};}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{variants.png}
  \caption{Integrity Violation\apt{Needs a machine-drawn version, as do the rest. Use colors consistently. No tags!}}
  \label{fig:variant}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{variants2.png}
  \caption{Internal Confidentiality Violation}
  \label{fig:variant2}
\end{figure}

@@

To illustrate the issues, note that {\tt f} has also stores a 0 below the stack pointer.
%
\begin{figure}
  \includegraphics[width=\columnwidth]{variants3.png}
  \caption{Return-time Confidentiality Violation}
  \label{fig:variant3}
\end{figure}
%
Now consider three elements: the address \(\SP - 4\), the address \(\SP + 12\),
and the register {\tt a0}. The execution from the right-hand state to its return
has changed the value at \(\SP - 4\), but that value matches that of the
left-hand variant at its return. Therefore, the change does not represent a leak.

>>>>

Of the mechanisms above, many are fundamentally ill-suited for offering formal guarantees,
as they aim to impede attackers but cannot provide universal protection. Shadow stacks,
for instance, aim to ``restrict the flexibility available in creating gadget chains''
\cite{Shanbhogue+19}, not to categorically
rule out attacks. Other mechanisms such as Softbound\cite{NagarakatteZMZ09} and Code-pointer
integrity\cite{Kuznetsov+14} are stronger, but most do not give formal guarantees.
Indeed, the secure calling conventions of Skorstengaard et al. \cite{SkorstengaardSTKJFP} and
Georges et al. \cite{Georges22:TempsDesCerises} is so far the sole line of work
to offer stack safety formalizations.

@@

\paragraph{Organization}

In the next section, we will give a brief overview of the structure of properties.
Then in \cref{sec:example}, we will walk through a function
call in a simple example machine and discuss how each of our properties applies to it,
informally. In the process we discuss some motivation for the properties, from a security
perspective.

In \cref{sec:machine} we formalize the machine model, 
its {\em security semantics}, and the stack safety properties built on these. 
\Cref{sec:enforcement} describes the micro-policies that we test,
\cref{sec:test} the testing framework itself, and
\cref{sec:relwork,sec:future} discuss related and future work.

%We consider a simple attacker, and discuss the harm that it can do to
%its caller. This perspective is essential to understanding how lazy enforcement can
%be safe: even if an attacker can overwrite some memory, if it cannot accomplish
%subvert the observable behavior of the caller, the system is safe.

\section{Building a Property}
\label{sec:ideas}

@@

\newcommand{\textbox}[1][]{\tikz \filldraw[fill=lightgray] (0,0) rectangle (30px,30px) node[pos=.5]{#1};}
\newcommand{\numbox}[1][]{\tikz \filldraw[fill=lightgray] (0,0) rectangle (20px,20px) node[pos=.5]{#1};}
\newcommand{\leftbox}[1][]{\tikz \filldraw[fill=blue!20] (0,0) rectangle (20px,20px) node[pos=.5]{#1};}
\newcommand{\rightbox}[1][]{\tikz \filldraw[fill=orange!20] (0,0) rectangle (20px,20px) node[pos=.5]{#1};}

@@

\begin{figure}
  \centering
  \[
  \stackrel{\texttt{res}}{\numbox[0]}
  \stackrel{\texttt{sens}}{\numbox[0]}
  \stackrel{\texttt{sec}}{\numbox[5]}
  \stackrel{\texttt{ra}}{\numbox[0]\numbox[0]}\]
  %
  \[\big\Downarrow\]
  %
  \[
  \stackrel{\texttt{res}}{\numbox[0]}
  \stackrel{\texttt{sens}}{\numbox[\bf 42]}
  \stackrel{\texttt{sec}}{\numbox[5]}
  \stackrel{\texttt{ra}}{\numbox[0]\numbox[0]}\]
  %
  \[\overbrace{
    \stackrel{\texttt{res}}{\numbox[0]}
    \stackrel{\texttt{sens}}{\numbox[\bf 42]}
    \stackrel{\texttt{sec}}{\numbox[5]}
    \stackrel{\texttt{ra}}{\numbox[0]\numbox[0]}
    \hspace{1cm}
    \stackrel{\texttt{res}}{\numbox[0]}
    \stackrel{\texttt{sens}}{\numbox[\bf 0]}
    \stackrel{\texttt{sec}}{\numbox[5]}
    \stackrel{\texttt{ra}}{\numbox[0]\numbox[0]}}\]
  \[\large\hookrightarrow 5 \hspace{1cm} \large\hookrightarrow 1\]

  \caption{Integrity Violation}
  \label{fig:variant}
\end{figure}

@@

\begin{figure}
    \centering
    \[
    \stackrel{\texttt{res}}{\numbox[0]}
    \stackrel{\texttt{sens}}{\numbox[0]}
    \stackrel{\texttt{sec}}{\numbox[5]}
    \stackrel{\texttt{ra}}{\numbox[0]\numbox[0]}\]
    %
    \[\overbrace{
    \stackrel{\texttt{res}}{\leftbox[0]}
    \stackrel{\texttt{sens}}{\leftbox[0]}
    \stackrel{\texttt{sec}}{\leftbox[5]}
    \stackrel{\texttt{ra}}{\leftbox[0]\leftbox[0]}
    \hspace{1cm}
    %
    \stackrel{\texttt{res}}{\rightbox[2]}
    \stackrel{\texttt{sens}}{\rightbox[3]}
    \stackrel{\texttt{sec}}{\rightbox[4]}
    \stackrel{\texttt{ra}}{\rightbox[5]\rightbox[6]}}
    \]
    \[\large\Downarrow \hspace{1cm} \large\hookrightarrow 5 \hspace{0.5cm} \not \approx \hspace{0.5cm}
    \large\hookrightarrow 4 \hspace{1cm} \large\Downarrow\]
    \[
    \stackrel{\texttt{res}}{\leftbox[\textbf{0}]}
    \stackrel{\texttt{sens}}{\leftbox[0]}
    \stackrel{\texttt{sec}}{\leftbox[5]}
    \stackrel{\texttt{ra}}{\leftbox[0]\leftbox[0]}
    \hspace{1cm}
    %
    \stackrel{\texttt{res}}{\rightbox[\textbf{2}]}
    \stackrel{\texttt{sens}}{\rightbox[3]}
    \stackrel{\texttt{sec}}{\rightbox[4]}
    \stackrel{\texttt{ra}}{\rightbox[5]\rightbox[6]}
    \]

  \caption{Internal Confidentiality Violation}
  \label{fig:variant2}
\end{figure}

@@

To illustrate the issues, note that {\tt f} has also stores a 0 below the stack pointer.
%
\begin{figure}
    \centering
    \[
    \stackrel{\texttt{res}}{\numbox[0]}
    \stackrel{\texttt{sens}}{\numbox[0]}
    \stackrel{\texttt{sec}}{\numbox[5]}
    \stackrel{\texttt{ra}}{\numbox[0]\numbox[0]}\]
    %
    \[\overbrace{
    \stackrel{\texttt{res}}{\leftbox[0]}
    \stackrel{\texttt{sens}}{\leftbox[0]}
    \stackrel{\texttt{sec}}{\leftbox[5]}
    \stackrel{\texttt{ra}}{\leftbox[0]\leftbox[0]}
    \hspace{1cm}
    %
    \stackrel{\texttt{res}}{\rightbox[2]}
    \stackrel{\texttt{sens}}{\rightbox[3]}
    \stackrel{\texttt{sec}}{\rightbox[4]}
    \stackrel{\texttt{ra}}{\rightbox[5]\rightbox[6]}}
    \]
    \[\large\Downarrow \hspace{1cm} \mathtt{a0} \leftbox[0] \hspace{1cm}
    \rightbox[7] \mathtt{a0} \hspace{1cm} \large\Downarrow\]
    \[
    \stackrel{\texttt{res}}{\leftbox[0]}
    \stackrel{\texttt{sens}}{\leftbox[0]}
    \stackrel{\texttt{sec}}{\leftbox[5]}
    \stackrel{\texttt{ra}}{\leftbox[0]\leftbox[0]}
    \hspace{1cm}
    %
    \stackrel{\texttt{res}}{\rightbox[2]}
    \stackrel{\texttt{sens}}{\rightbox[3]}
    \stackrel{\texttt{sec}}{\rightbox[4]}
    \stackrel{\texttt{ra}}{\rightbox[5]\rightbox[6]}
    \]
    \[\mathtt{a0} \leftbox[0] \hspace{1cm}
    \mathtt{a0} \rightbox[0]\]
  \caption{Return-time Confidentiality Violation}
  \label{fig:variant3}
\end{figure}
%
Now consider three elements: the address \(\SP - 4\), the address \(\SP + 12\),
and the register {\tt a0}. The execution from the right-hand state to its return
has changed the value at \(\SP - 4\), but that value matches that of the
left-hand variant at its return. Therefore, the change does not represent a leak.
